Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm

### 2011-07-14 ltx/data_paper/DataPaper.tex
added computational complexity initialisation and stopping

By computing $\boldsymbol\Xi$-variables, calculation of the lower bound is complete. The lower bound now can be used in M-step to find a set of new parameters. An approximate for computational complexity of the algorithm is given in Table~\ref{table:ComputationalComplexity}. Note that computational complexity of the nonlinear terms depend on the number of points, $n_r$ used for spatial discritisation.  In this table only the computational complexity of variables specific to the presented algorithm are shown, and the computational complexity of the URTSS can be found elsewhere \cite{Sarkka2010}.

@@

%  Increasing the joint likelihood function increases the likelihood function $p(\mathbf Y;\boldsymbol \Theta)$ \cite{McLachlan1997}. 
 \begin {table}
\scalebox{1}{\begin{tabular}{lll}
\hline \hline
Variable &Equation&Order\\ \hline\\
$\mathbf q(\mathbf{x}_t)$&\eqref{}&$O(n_xn_{\theta}n_r)$\\
$\boldsymbol\Xi_0, \boldsymbol\Xi_1, \boldsymbol\Xi_2$&\eqref{eq:Xi0Gibson}, \eqref{eq:Xi1Gibson}, \eqref{eq:Xi2Gibson}&$O(Tn_x^2)$\\
$\boldsymbol\Xi_3, \boldsymbol\Xi_5$&\eqref{eq:Xi3Ap},\eqref{eq:Xi5Ap} &$O(Tn_x^3)+O(Tn_x^2n_r)+O(Tn_xn_{\theta}n_r)$ \\
$\boldsymbol\Xi_4$&\eqref{eq:Xi4Ap}&$O(Tn_x^4n_{\theta})+O(Tn_x^3n_{\theta}^2)+O(Tn_x^3n_{\theta}n_r)$\\
$\hat{\xi}$&\eqref{eq:xihat}&$O(n_x^3+n_{\theta}^3)$\\
$\hat{\boldsymbol\theta} $&\eqref{eq:thetahat}&$O(n_{\theta}^3)$\\
$\hat{\sigma}_{\epsilon}^2$&\eqref{eq:sigmaepsilonhat}&$O(Tn_x^2n_y)+O(Tn_xn_y^2)$\\
$\hat{\sigma}_{d}^2$&\eqref{eq:sigmadhat}&$O(n_x^3)+O(n_{\theta}^2)$\\
\hline \hline
\end{tabular}}
 \caption {Computational complexity of EM algorithm} 
\label{table:ComputationalComplexity}
\end {table}
\section{Results}
This section demonstrates the identification of the proposed IDE model using simulated data. The observation noise and the field disturbance were unknown to the estimation algorithm and no assumption were made on the shape of the connectivity kernel. Data was generated employing state-space equation (\ref{eq:StateEq}-\ref{eq:QmatrixForSigmapoints}) using a square grid of $n_x= 9 \times 9$ equally spaced ($2.5~mm$) basis functions with $\sigma_{\phi}^2=1.58~mm^2$, assuming the observed field can be fully described by the decomposed IDE model. 100 realizations of 500 ms of data were generated and the first 100 ms was discarded allowing the model's dynamic to reach steady state. Observation noise variance was set to 0.2. All other parameters for the simulation were the same as Chapter 2 (see Table 1.1).

>>>>

By computing $\boldsymbol\Xi$-variables, calculation of the lower bound is complete. The lower bound now can be used in M-step to find a set of new parameters. 

@@

\subsection{Initialisation, stopping and algorithmic complexity}
For initialisation a bounded random state sequence is employed to ensure that the initial parameter estimates describe a stable kernel, although the quality of such initial estimate is poor, it provides an initialisation that typically ensues a satisfactory convergence of the algorithm, as illustrated numerically in Result section. The stopping criterion is usually associated with either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}. Here the change in parameter estimates, $L^2$-norm in the case of connectivity kernel weights, is used to halt the excecution of the algorithm. Besides, the lower bound on the likelihood function is also monitored to further investigate the convergence of the EM algorithm.

An approximate for computational complexity of the algorithm is given in Table~\ref{table:ComputationalComplexity}.  In this table only the computational complexity of variables specific to the presented algorithm are shown, and the computational complexity of the URTSS can be found in \cite{Sarkka2010}. Note that computational complexity of the nonlinear terms depend on the number of points, $n_r$ used for spatial discritisation. Calculation of the nonlinear terms $\boldsymbol\Xi$ is clearly the slowest part of the algorithm where $n_x$ is the problematic dimension, more specifically the matrix $\boldsymbol\Xi_4 $ has complexity that grows with the order of four in $n_x$. Although number of states  is decoupled from number of observation locations (recording electrods) , and the number of basis functions required for the connectivity kernel decomposition, it does tend to grow linearly in one dimensional space and quadratically in two dimensional space. This growth assumes that the basis functions are arragend regularly in space, which is not a necessary construction.

%  Increasing the joint likelihood function increases the likelihood function $p(\mathbf Y;\boldsymbol \Theta)$ \cite{McLachlan1997}. 
 \begin {table}
\begin{center}
\scalebox{1}{\begin{tabular}{lll}
\hline \hline
Variable &Equation&Order\\ \hline\\
$\mathbf q(\mathbf{x}_t)$&\eqref{}&$O(n_xn_{\theta}n_r)$\\
$\boldsymbol\Xi_0, \boldsymbol\Xi_1, \boldsymbol\Xi_2$&\eqref{eq:Xi0Gibson}, \eqref{eq:Xi1Gibson}, \eqref{eq:Xi2Gibson}&$O(Tn_x^2)$\\
$\boldsymbol\Xi_3, \boldsymbol\Xi_5$&\eqref{eq:Xi3Ap},\eqref{eq:Xi5Ap} &$O(Tn_x^3)+O(Tn_x^2n_r)+O(Tn_xn_{\theta}n_r)$ \\
$\boldsymbol\Xi_4$&\eqref{eq:Xi4Ap}&$O(Tn_x^4n_{\theta})+O(Tn_x^3n_{\theta}^2)+O(Tn_x^3n_{\theta}n_r)$\\
$\hat{\xi}$&\eqref{eq:xihat}&$O(n_x^3+n_{\theta}^3)$\\
$\hat{\boldsymbol\theta} $&\eqref{eq:thetahat}&$O(n_{\theta}^3)$\\
$\hat{\sigma}_{\epsilon}^2$&\eqref{eq:sigmaepsilonhat}&$O(Tn_x^2n_y)+O(Tn_xn_y^2)$\\
$\hat{\sigma}_{d}^2$&\eqref{eq:sigmadhat}&$O(n_x^3)+O(n_{\theta}^2)$\\
\hline \hline
\end{tabular}}
 \caption {Computational complexity of EM algorithm} 
\label{table:ComputationalComplexity}
\end{center}
\end {table}
\section{Results}
This section demonstrates the identification of the proposed IDE model using simulated data. The observation noise and the field disturbance were unknown to the estimation algorithm and no assumption were made on the shape of the connectivity kernel. Data was generated employing state-space equation (\ref{eq:StateEq}-\ref{eq:QmatrixForSigmapoints}) using a square grid of $n_x= 9 \times 9$ equally spaced ($2.5~mm$) basis functions with $\sigma_{\phi}^2=1.58~mm^2$, assuming the observed field can be fully described by the decomposed IDE model. 100 realizations of 500 ms of data were generated and the first 100 ms was discarded allowing the model's dynamic to reach steady state. Observation noise variance was set to 0.2. All other parameters for the simulation were the same as Chapter 2 (see Table 1.1).
